---
title: "Assignment 3"
subtitle: "Almond Profit"
format: html
author: "Nicole Pepper"
editor_options: 
  chunk_output_type: console
---

### Set up Workspace 
```{r}
# Load libraries
library(here)
library(tidyverse)
library(dplyr)
```

```{r}
# Read in climate data
clim_data <- read.table(here("HW2","data","clim.txt"), header = TRUE)
```

```{r}
# ---- Prep clim data for model ----

# Group clim data by month
monthly_clim <- clim_data |>
  group_by(year,month) %>%
  summarise(
    tmin_avg = mean(tmin_c,na.rm = TRUE),
    precip_total = sum(precip, na.rm = TRUE),
    .groups = "drop")

# Filter clim data for avg temp in Feb
Temp_Feb <- monthly_clim |>
  filter(month == 2) |>
  select(year,
         Temp_Feb = tmin_avg)

# Filter clim data for total precip in Jan
Precip_Jan <- monthly_clim |>
  filter(month == 1) |>
  select(year,
         Precip_Jan = precip_total)

# Join filtered clim data back together
clim_data_clean <- inner_join(Temp_Feb, Precip_Jan, by = "year")

```


### 1. Develop a profit model for your almond yield

Use your imagination to think about what would make sense
Make sure you include parameters
You could assume a baseline profit and then adjust according to the anomaly


```{r}
# Source profit model
source(here("HW3","R","almond_profit_func.R"))
```


### 2. Apply the profit model to your almond yield model

There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function)


```{r}
# Apply profit model to each year
almond_profits <- clim_data_clean |>
  mutate(profit = mapply(
    almond_profit,
    Temp_Feb = Temp_Feb,
    Precip_Jan = Precip_Jan))

# Check out result
head(almond_profits)

```

### 3. Perform  a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

```{r}

# ---- Perform sensitivity analysis on temp and precip ----

# Create grid of temp and precip range
temp_seq <- seq(5, 15, by = 0.25)
precip_seq <- seq(0, 150, by = 5)

# Define grid
clim_grid <- expand.grid(
  Temp_Feb = temp_seq,
  Precip_Jan = precip_seq)

# Test sensitivity to temp and precip
clim_grid$profit <- mapply(
  almond_profit,
  Temp_Feb = clim_grid$Temp_Feb,
  Precip_Jan = clim_grid$Precip_Jan)

# Apply profit model to all years
almond_profits <- clim_data_clean |>
  mutate(profit = mapply(
    almond_profit,
    Temp_Feb = Temp_Feb,
    Precip_Jan = Precip_Jan))

 

```


### 4. Create 2 graphs
    - one that shows yield anomaly for each year, accounting for uncertainty in the parameters
```{r}
n_sim <- 100

# Create grid of temp and precip range
temp_sims <- rnorm(n_sim, mean = clim_data_clean$Temp_Feb, sd = 1)
precip_sims <- rnorm(n_sim, mean = clim_data_clean$Precip_Jan, sd = 20)

# Apply to 100 simulations
yield_uncertainty <- clim_data_clean |>
  slice(rep(1:n(), each = n_sim)) |>
  mutate(
    Temp_Feb_sim = rnorm(n(), Temp_Feb, sd = 1),
    Precip_Jan_sim = rnorm(n(), Precip_Jan, sd = 20),
    yield_anomaly = almond_yield(
      Temp_Feb_sim, Precip_Jan_sim,
      T_Coef1 = -0.015, T_Coef2 = -0.0046,
      P_Coef1 = -0.07, P_Coef2 = 0.0043, I = 0.28))

# Plot yield anomaly for each year accounting for uncertainty in params (temp and precip)
ggplot(yield_uncertainty, aes(x = year, y = yield_anomaly, group = year)) +
  geom_boxplot(fill = "cornflowerblue") +
  labs(
    title = "Yield Anomaly by Year with Uncertainty For Temp & Precip",
    x = "Year",
    y = "Yield Anomaly (tons/acre)") +
  theme_minimal()
```
  
- one that show how yield anomaly in general varies with your parameters
```{r}

# ---- Visualize sensitivity to temp ----

# Define range for temp
temp_seq <- seq(5, 15, by = 0.1)

# Define df with precip constant
yield_temp <- tibble(
  Temp_Feb = temp_seq,
  Precip_Jan = 50)

# Calculate yield anomaly with temp varied
yield_temp$yield_anomaly <- almond_yield(
  Temp_Feb = yield_temp$Temp_Feb,
  Precip_Jan = yield_temp$Precip_Jan,
  T_Coef1 = -0.015,
  T_Coef2 = -0.0046,
  P_Coef1 = -0.07,
  P_Coef2 = 0.0043,
  I = 0.28)

# Plot how yield anomaly varies with temp
ggplot(yield_temp, aes(x = Temp_Feb, y = yield_anomaly)) +
  geom_line(color = "cornflowerblue", linewidth = 1) +
  labs(
    title = "Yield Anomaly Sensitivity to Temperature",
    x = "February Temperature (°C)",
    y = "Yield Anomaly (tons/acre)") +
  theme_minimal()


```

```{r}
# ---- Visualize sensitivity to precip ----

# Define range for precip
precip_seq <- seq(0, 150, by = 1)

# Define df with temp constant
yield_precip <- tibble(
  Precip_Jan = precip_seq,
  Temp_Feb = 10)

# Calculate yield anomaly with precip varied
yield_precip$yield_anomaly <- almond_yield(
  Temp_Feb = yield_precip$Temp_Feb,
  Precip_Jan = yield_precip$Precip_Jan,
  T_Coef1 = -0.015,
  T_Coef2 = -0.0046,
  P_Coef1 = -0.07,
  P_Coef2 = 0.0043,
  I = 0.28)

# Plot how yield anomaly varies with precip
ggplot(yield_precip, aes(x = Precip_Jan, y = yield_anomaly)) +
  geom_line(color = "cornflowerblue", linewidth = 1) +
  labs(
    title = "Yield Anomaly Sensitivity to Precipitation",
    x = "January Precipitation (mm)",
    y = "Yield Anomaly (tons/acre)") +
  theme_minimal()
```

### 5. Write a short paragraph (in a Quatro document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

- Variations in temperature and precipitation can have significant impacts on model performance.
- Yield anomaly seems to decrease as temperature increases.
- Yield anomaly seems to increase as precipitation increases.
